---
title: "Load & Review Data v2 (get all Morgue Donors sorted)"
author: "Wesley"
date: "2022/05/03"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load all of the Morgue Data

To start, we need to load all of the Excel files. To do this, we will use the
readxl package. In addition, we're going to need to manipulate strings in 
the following, so let's pre-emptively load that package.

```{r libraries, error = FALSE, message = FALSE, warning = FALSE}
library("readxl")
library("stringr")
```

### Start with M_DONOR_1, as Template

Foot 1 has many sheets, each of which is a nicely organized listing of VOCs
as detected by the GCxGC system's built-in scanning software, and exported. 
We will load them all. 

```{r load_data}
sheets <- excel_sheets("MORGUE/M_DONOR_1.xlsx")  # note: spaces in names
M_DONOR_1 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "MORGUE/M_DONOR_1.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(M_DONOR_1) <- sheets
```


Having things organized by sample is a good thing, but we really will
find it annoying finding things long-term. In the following, I will take each
of the lists of compounds relevant to Foot 1, and merge them into a single
data.frame object, so we can then filter that frame down to get a specific
compound, or a specific day/observation, and so on.

```{r merge_sheets}
M_DONOR_1m <- do.call("rbind", M_DONOR_1)
names(M_DONOR_1m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs","1st Dimension Time", "2nd Dimension Time", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

A quick print of the data structure shows that the RTs, QuantMass and BaseMass
won't be usable as numbers. The RTs need to be split, as well.

```{r split_RTs}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_1m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_1m$RTs, ","), FUN = "[[", 2)
         )
       )
M_DONOR_1m <- data.frame(M_DONOR_1m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples}
unique(M_DONOR_1m$Sample)
save(file = "dat/M_DONOR_1.rda", M_DONOR_1m)
```

### Now Do M_DONOR_2

```{r load_data2}
sheets <- excel_sheets("MORGUE/M_DONOR_2.xlsx")  # note: spaces in names
M_DONOR_2 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "MORGUE/M_DONOR_2.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(M_DONOR_2) <- sheets

M_DONOR_2m <- do.call("rbind", M_DONOR_2)
names(M_DONOR_2m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs","1st Dimension Time", "2nd Dimension Time", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs2}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_2m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_2m$RTs, ","), FUN = "[[", 2)
         )
       )
M_DONOR_2m <- data.frame(M_DONOR_2m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(M_DONOR_2m$Sample)
save(file = "dat/M_DONOR_2.rda", M_DONOR_2m)
```

### Now Do M_DONOR_3

```{r load_data3}
sheets <- excel_sheets("MORGUE/M_DONOR_3.xlsx")  # note: spaces in names
M_DONOR_3 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "MORGUE/M_DONOR_3.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(M_DONOR_3) <- sheets

M_DONOR_3m <- do.call("rbind", M_DONOR_3)
names(M_DONOR_3m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "1st Dimension Time", "2nd Dimension Time", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs3}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_3m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_3m$RTs, ","), FUN = "[[", 2)
         )
       )
M_DONOR_3m <- data.frame(M_DONOR_3m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples3}
unique(M_DONOR_3m$Sample)
save(file = "dat/M_DONOR_3.rda", M_DONOR_3m)
```

### Now Do M_DONOR_4

```{r load_data4}
sheets <- excel_sheets("MORGUE/M_DONOR_4.xlsx")  # note: spaces in names
M_DONOR_4 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "MORGUE/M_DONOR_4.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(M_DONOR_4) <- sheets

M_DONOR_4m <- do.call("rbind", M_DONOR_4)
names(M_DONOR_4m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "1st Dimension Time", "2nd Dimension Time", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_4m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_4m$RTs, ","), FUN = "[[", 2)
         )
       )
M_DONOR_4m <- data.frame(M_DONOR_4m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples4}
unique(M_DONOR_4m$Sample)
save(file = "dat/M_DONOR_4.rda", M_DONOR_4m)
```

### Now Do M_DONOR_6

```{r load_data5}
sheets <- excel_sheets("MORGUE/M_DONOR_6.xlsx")  # note: spaces in names
M_DONOR_6 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "MORGUE/M_DONOR_6.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(M_DONOR_6) <- sheets

M_DONOR_6m <- do.call("rbind", M_DONOR_6)
names(M_DONOR_6m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs","1st Dimension Time", "2nd Dimension Time", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs6}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_6m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_6m$RTs, ","), FUN = "[[", 2)
         )
       )
M_DONOR_6m <- data.frame(M_DONOR_6m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples6}
unique(M_DONOR_6m$Sample)
save(file = "dat/M_DONOR_6.rda", M_DONOR_6m)
```

### Now Do M_DONOR_7

```{r load_data6}
sheets <- excel_sheets("MORGUE/M_DONOR_7.xlsx")  # note: spaces in names
M_DONOR_7 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "MORGUE/M_DONOR_7.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(M_DONOR_7) <- sheets

M_DONOR_7m <- do.call("rbind", M_DONOR_7)
names(M_DONOR_7m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs","1st Dimension Time", "2nd Dimension Time", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs7}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_7m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_7m$RTs, ","), FUN = "[[", 2)
         )
       )
M_DONOR_7m <- data.frame(M_DONOR_7m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples7}
unique(M_DONOR_7m$Sample)
save(file = "dat/M_DONOR_7", M_DONOR_7m)
```

### Now Do M_DONOR_8

```{r load_data7}
sheets <- excel_sheets("MORGUE/M_DONOR_8.xlsx")  # note: spaces in names
M_DONOR_8 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "MORGUE/M_DONOR_8.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(M_DONOR_8) <- sheets

M_DONOR_8m <- do.call("rbind", M_DONOR_8)
names(M_DONOR_8m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "1st Dimension Time", "2nd Dimension Time", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs8}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_8m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_8m$RTs, ","), FUN = "[[", 2)
         )
       )
M_DONOR_8m <- data.frame(M_DONOR_8m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples8}
unique(M_DONOR_8m$Sample)
save(file = "dat/M_DONOR_8.rda", M_DONOR_8m)
```

### Now Do M_DONOR_9

```{r load_data8}
sheets <- excel_sheets("MORGUE/M_DONOR_9.xlsx")  # note: spaces in names
M_DONOR_9 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "MORGUE/M_DONOR_9.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(M_DONOR_9) <- sheets

M_DONOR_9m <- do.call("rbind", M_DONOR_9)
names(M_DONOR_9m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs","1st Dimension Time","2nd Dimension Time", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r}
#M_DONOR_9_repl
#M_DONOR_9_cont
#M_DONOR_9 <- c(M_DONOR_9_repl, M_DONOR_9_contr)   # before the do.call("rbind") line
```


```{r split_RTs9}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_9m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(M_DONOR_9m$RTs, ","), FUN = "[[", 2)
         )
       )
M_DONOR_9m <- data.frame(M_DONOR_9m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples9}
unique(M_DONOR_9m$Sample)
save(file = "dat/M_DONOR_9.rda", M_DONOR_9m)
```


