---
title: "Load & Review Data v2 (get all REST donors sorted)"
author: "Wesley"
date: "2022/03/28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load all of the REST Data

To start, we need to load all of the Excel files. To do this, we will use the
readxl package. In addition, we're going to need to manipulate strings in 
the following, so let's pre-emptively load that package.

```{r libraries, error = FALSE, message = FALSE, warning = FALSE}
library("readxl")
library("stringr")
```

### Start with REST donor 1, as Template

Foot 1 has many sheets, each of which is a nicely organized listing of VOCs
as detected by the GCxGC system's built-in scanning software, and exported. 
We will load them all. 

```{r load_data}
sheets <- excel_sheets("/home/wburr/Darshil_Compounds/REST/RH1new.xlsx")  # note: spaces in names
RH1new <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Darshil_Compounds/REST/RH1new.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(RH1new) <- sheets
```

Having things organized by sample is a good thing, but we really will
find it annoying finding things long-term. In the following, I will take each
of the lists of compounds relevant to Foot 1, and merge them into a single
data.frame object, so we can then filter that frame down to get a specific
compound, or a specific day/observation, and so on.

Before we do this, the first sheets has the observations duplicated across
the columns (wups!). Removing that:

```{r remove_sheet1}
RH1new[[1]] <- RH1new[[1]][, 1:ncol(RH1new[[2]])]
names(RH1new[[1]]) <- names(RH1new[[2]])  # fix names of sheet 1
```

```{r merge_sheets}
RH1newm <- do.call("rbind", RH1new)
names(RH1newm) <- c("Peak#","PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN")
```

A quick print of the data structure shows that the RTs, QuantMass and BaseMass
won't be usable as numbers. The RTs need to be split, as well.

```{r split_RTs}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(RH1newm$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(RH1newm$RTs, ","), FUN = "[[", 2)
         )
       )
RH1newm <- data.frame(RH1newm[, c(1:5, 6:13)], RT1 = RT1, RT2 = RT2)
```

```{r samples}
unique(RH1newm$Sample)
save(file = "/home/wburr/Darshil_Compounds/REST/RH1new2.rda", RH1newm)
```

### Now Do REST donor 2

```{r load_data2}
sheets <- excel_sheets("/home/wburr/Darshil_Compounds/REST/RH2.xlsx")  # note: spaces in names
RH2 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Darshil_Compounds/REST/RH2.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(RH2) <- sheets

RH2m <- do.call("rbind", RH2)
names(RH2m) <- c("Peak#","PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN")
                   
```

```{r split_RTs2}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(RH2m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(RH2m$RTs, ","), FUN = "[[", 2)
         )
       )
RH2m <- data.frame(RH2m[, c(1:5, 6:13)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(RH2m$Sample)
save(file = "/home/wburr/Darshil_Compounds/REST/RH2.rda", RH2m)
```

### Now Do Donor 3

```{r load_data3}
sheets <- excel_sheets("/home/wburr/Darshil_Compounds/REST/RH3N1.xlsx")  # note: spaces in names
RH3N1 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Darshil_Compounds/REST/RH3N1.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(RH3N1) <- sheets

RH3N1m <- do.call("rbind", RH3N1)
names(RH3N1m) <- c("Peak#","Peaknumber", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN")

```

```{r split_RTs3}
# filter out the missing RTs/names
RH3N1m <- RH3N1m[!is.na(RH3N1m$Name), ]
# which-split
which_split <- unlist(lapply(strsplit(RH3N1m$RTs, ","), length))
RH3N1mA <- RH3N1m[which_split == 1, ]
RH3N1mB <- RH3N1m[which_split == 2, ]

# the RTs that can be split, split
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(RH3N1mB$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(RH3N1mB$RTs, ","), FUN = "[[", 2)
         )
       )
# merge them. The 1mA is probably garbage anyway, but it's kept at this stage JIC
RH3N1mB <- data.frame(RH3N1mB[, c(1:5, 6:13)], RT1 = RT1, RT2 = RT2)
RH3N1mA <- data.frame(RH3N1mA[, c(1:5, 6:13)], RT1 = NA, RT2 = NA)
RH3N1m <- rbind(RH3N1mA, RH3N1mB)
```

```{r samples3}
unique(RH3N1m$Sample)
save(file = "/home/wburr/Darshil_Compounds/REST/RH3N1.rda", RH3N1m)
```

### Now Do REST donor 4

```{r load_data4}
sheets <- excel_sheets("/home/wburr/Darshil_Compounds/REST/RH4.xlsx")  # note: spaces in names
RH4 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Darshil_Compounds/REST/RH4.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(RH4) <- sheets

RH4m <- do.call("rbind", RH4)
names(RH4m) <- c("Peak#","PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN")

```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(RH4m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(RH4m$RTs, ","), FUN = "[[", 2)
         )
       )
RH4m <- data.frame(RH4m[, c(1:5, 6:13)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(RH4m$Sample)
save(file = "/home/wburr/Darshil_Compounds/REST/RH4.rda", RH4m)
```

### Now Do REST donor 6

```{r load_data5}
sheets <- excel_sheets("/home/wburr/Darshil_Compounds/REST/RH6.xlsx")  # note: spaces in names
RH6 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Darshil_Compounds/REST/RH6.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(RH6) <- sheets

RH6m <- do.call("rbind", RH6)
names(RH6m) <- c("Peak#","PeakNumber", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(RH6m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(RH6m$RTs, ","), FUN = "[[", 2)
         )
       )
RH6m <- data.frame(RH6m[, c(1:5, 6:13)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(RH6m$Sample)
save(file = "/home/wburr/Darshil_Compounds/REST/RH6.rda", RH6m)
```



### Now Do Donor 7

```{r load_data7}
sheets <- excel_sheets("/home/wburr/Darshil_Compounds/REST/RH7.xlsx")  # note: spaces in names
RH7 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Darshil_Compounds/REST/RH7.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(RH7) <- sheets

RH7m <- do.call("rbind", RH7)
names(RH7m) <- c("Peak#","PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(RH7m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(RH7m$RTs, ","), FUN = "[[", 2)
         )
       )
RH7m <- data.frame(RH7m[, c(1:5, 6:13)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(RH7m$Sample)
save(file = "/home/wburr/Darshil_Compounds/REST/RH7.rda", RH7m)
```

### Now Do Donor 8

```{r load_data8}
sheets <- excel_sheets("/home/wburr/Darshil_Compounds/REST/RH8.xlsx")  # note: spaces in names
RH8 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Darshil_Compounds/REST/RH8.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(RH8) <- sheets

RH8m <- do.call("rbind", RH8)
names(RH8m) <- c("Peak#","PeakNum", "Sample", "Name", "Formula", "RTs", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(RH8m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(RH8m$RTs, ","), FUN = "[[", 2)
         )
       )
RH8m <- data.frame(RH8m[, c(1:5, 6:13)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(RH8m$Sample)
save(file = "/home/wburr/Darshil_Compounds/REST/RH8.rda", RH8m)
```

### Now Do Donor 9 

```{r load_data9}
sheets <- excel_sheets("/home/wburr/Darshil_Compounds/REST/RH9.xlsx")  # note: spaces in names
RH9  <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Darshil_Compounds/REST/RH9.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(RH9) <- sheets

RH9m <- do.call("rbind", RH9)
names(RH9m) <- c( "Peak #", "Peak Number", "Sample",	"Name",	"Formula", "RTs", "Similarity", "Area",	                             "Height",	"Quant Masses", "Base Mass", "Quant S/N", "Peak S/N")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(RH9m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(RH9m$RTs, ","), FUN = "[[", 2)
         )
       )
RH9m <- data.frame(RH9m[, c(1:5, 6:13)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(RH9m$Sample)
save(file = "/home/wburr/Darshil_Compounds/dat/RH9.rda", RH9m)
```

